Greg Caporaso has been an ideal advisor, providing countless ideas
and opportunities, along with the support, time, and encouragement I needed to
try, fail, learn, and grow as a developer, thinker, and individual.

Matt Dillon has served as both teacher and role model, modeling ways to learn,
to lead, and to live one's values. He has impressed upon me the importance of
asking the right questions, investing in knowledge, standing up when you get
knocked down, and recognizing when the juice is just not worth the squeeze.

Evan Bolyen taught me what a stack frame was, after two years at University had
failed to do so. His advice on the importance of ambition and failure to
personal and professional growth has been absolutely invaluable, and his unique
combination of empathy, erudition, and fascination with the world around him is
a constant inspiration.

Anthony Simard, Emily Borsom, and all the other members of the Caporaso and Cope labs have
been the best friends, collaborators, and listening ears anyone could hope for.
My committee, Dr. Marco Gerosa and Dr. Nicholas Bokulich, have provided invaluable
advice and critique.

Heidi Rodenhizer, my wife and partner, has shown me what it is to work happily
towards a goal, and has supported me through all of the ups and downs of grad
school while a PhD student herself. My parents, Cecilia Roehl and Bill Keefe,
and my sister Kathleen, continue to teach, motivate, and inspire me every day.

This work was supported in part by the Arizona Alzheimer’s Research Center, Inc
(Award number CTR040636) to Emily Cope and Greg Caporaso, and by the National
Cancer Institute (Award number 1U24CA248454-01) to Greg Caporaso, with
additional support along the way from the Chan Zuckerberg Initiative, the
National Institutes of Health, and the National Science Foundation.