\chapter{Survey Fulltext}
\label{app:survey}

\noindent Unless otherwise noted, all questions were asked on a 7-value likert
scale, on which participants rate their level of agreement with a statement as
one of the following:
\begin{enumerate}
    \item Strongly Disagree
    \item Disagree
    \item Somewhat Disagree
    \item Neither Agree nor Disagree
    \item Somewhat Agree
    \item Agree
    \item Strongly Agree
\end{enumerate}

\hline

\subsection*{Career/Discipline}

\begin{itemize}
    \item What is your current occupation/career stage? \emph{[Non-likert: text field]}
    \item Please indicate your relevant fields or disciplines. \emph{[Non-likert: text field]}
\end{itemize}

\subsection*{Technology Use}

\noindent Please rate your level of agreement with the following statements:

\begin{itemize}
    \item I frequently use a command shell (e.g. through Terminal on MacOS or PowerShell on Windows).
    \item I feel comfortable using programming languages (e.g. R, Python, Java, etc.)
    \item I feel comfortable using QIIME 2.
    \item I feel comfortable using QIIME 2's provenance tracking features (e.g. automatic tracking of methods and citations, provenance graphs in q2view)
    \item I am satisfied with my current analysis workflow in QIIME 2 (i.e. how you analyze and report on your data).
    \item QIIME 2’s provenance tracking features are a useful tool in my current analysis workflow
\end{itemize}

\subsection*{Attitude}

\begin{itemize}
    \item I enjoy learning to use new computational tools
    \item It is generally easy to learn to use new computational tools
    \item I enjoy using QIIME 2
    \item It is generally easy for me to learn new things in QIIME 2
\end{itemize}


\subsection*{Perceived ease of use}

\noindent Imagine you have all of the software tools demonstrated today included in your
installation of QIIME 2. Please rate your level of agreement with the following
statements:
\begin{itemize}
    \item Generating a script for my preferred user interface from an existing QIIME 2 Result will be easy for me.
    \item Re-running a completed analysis in a different interface will be easy for me.
    \item Collecting all citations from a full analysis will be easy for me.
    \item Reading and learning from other users’ methods will be easy for me.
    \item Training new teammates in QIIME 2 analysis will be easy for me.
    \item QIIME 2 documentation is easy for me to access.
    \item QIIME 2 documentation is helpful for me when I use it.
    \item Overall, I believe these features are easy to use.
\end{itemize}

\subsection*{Perceived usefulness}


\noindent Again, imagine you have all of the software tools demonstrated today included in
your installation of QIIME 2.  Please rate your level of agreement with the
following statements:
\begin{itemize}
    \item The demonstrated features will help me reproduce and extend my own studies
    \item The demonstrated features will help me reproduce and extend other scientists’ studies.
    \item The demonstrated features will help me feel more confident in the validity of published QIIME 2 results.
    \item The demonstrated features will make training new teammates easier.
    \item The demonstrated features will make it easier for me to share knowledge with collaborators.
    \item The demonstrated features will help me learn new analysis techniques more efficiently.
    \item The demonstrated features will save me time and energy recording my analysis methods and parameters.
    \item The demonstrated features will make it easier for me to write a complete methods section.
    \item The demonstrated features will help me cite all of the sources relevant to my computational analysis.
    \item The demonstrated features will be useful in automating common or frequently-repeated analysis pipelines.
    \item The demonstrated features will improve my analysis workflow.
    \item These features are useful for me as a QIIME 2 user.
\end{itemize}

\subsection*{Net Promoter Score}

\noindent Considering your overall experience of these provenance replay features, how
likely would you be to recommend them to a friend or colleague who also uses
QIIME 2? \\
\emph{[Non-likert: Select integer value from 1-10]}

\subsection*{Value of features to the user}

\begin{itemize}
    \item Please select all features you expect to incorporate into your
    analysis workflow in QIIME 2 (i.e. how you analyze and report on your data).
    \emph{[Non-Likert: Select-many of demonstrated features]}
    \item Please rank the demonstrated features in order from most valuable to
    you, to least valuable to you. \emph{[Non-Likert: Rank]}
\end{itemize}
